---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

<!-- badges: start -->
[![R build status](https://github.com/tnagler/svines/workflows/R-CMD-check/badge.svg)](https://github.com/tnagler/svines/actions)
[![Codecov test coverage](https://codecov.io/gh/tnagler/svines/branch/master/graph/badge.svg)](https://codecov.io/gh/tnagler/svines?branch=master)
[![CRAN status](https://www.r-pkg.org/badges/version/svines)](https://CRAN.R-project.org/package=svines)
<!-- badges: end -->

An R package that provides functionality to fit and simulate from [stationary vine copula models 
for time series](https://github.com/tnagler/tnagler.github.io/raw/master/svines.pdf).

The package is build on top of [rvinecopulib](https://github.com/vinecopulib/rvinecopulib)
and [univariateML](https://github.com/JonasMoss/univariateML).


## Installation

Install the development version from Github.

``` r
# install.packages("devtools")
devtools::install_github("tnagler/svines")
```

## Usage

For detailed documentation and examples, see the [API documentation](https://tnagler.github.io/svines/).

``` {r}
library(svines)
data(returns)  # data set of stock returns
returns <- returns[1:500, 1:3]
```

### Fitting models

```{r}
fit <- svine(returns, p = 1)  # Markov order 1
summary(fit)
```

```{r}
contour(fit$copula)
```

### Simulation

`svine_sim()` can be used in two different ways:

#### Generate a new time series of length 500
```{r}
sim <- svine_sim(n = 500, rep = 1, model = fit)
pairs(sim)
pairs(returns)
```

#### Sample conditionally on the past given the past

```{r}
sim <- svine_sim(n = 1, rep = 100, model = fit, past = returns)
pairs(t(sim[1, , ]))
```


### Standard errors

```{r}
# compute standard errors for each parameter
se <- sqrt(diag(svine_avar(returns, fit)))
```
The standard errors correspond to (in this order): parameters of first margin,
parameters of second margin, ...., copula parameters in first tree (excluding 
duplicates), ...

The asymptotic variance is the basis of the bootstrap procedure in the paper.
To generate new models from the asymptotic distribution, use
```{r}
models <- svine_sim_se_models(2, fit)
summary(models[[1]])
summary(models[[1]])
```

