---
output: github_document
---

# svines

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  message = FALSE,
  warning = FALSE
)
set.seed(2026)
```

<!-- badges: start -->
[![R build status](https://github.com/tnagler/svines/workflows/R-CMD-check/badge.svg)](https://github.com/tnagler/svines/actions)
[![Codecov test coverage](https://codecov.io/gh/tnagler/svines/branch/main/graph/badge.svg)](https://app.codecov.io/gh/tnagler/svines?branch=main)
[![CRAN status](https://www.r-pkg.org/badges/version/svines)](https://CRAN.R-project.org/package=svines)
<!-- badges: end -->

An R package that provides functionality to fit and simulate from [stationary vine copula models 
for time series](https://arxiv.org/abs/2008.05990).

The package is built on top of [rvinecopulib](https://github.com/vinecopulib/rvinecopulib)
and [univariateML](https://github.com/JonasMoss/univariateML).


## Installation

Install the released version from CRAN.

``` r
install.packages("svines")
```

Install the development version from GitHub with `remotes`.

``` r
# install.packages("remotes")
remotes::install_github("tnagler/svines")
```

## Usage

For detailed documentation and examples, see the
[package website](https://tnagler.github.io/svines/).

Use `svine()` for observed data: it estimates the marginal distributions and
the S-vine copula. Use `svinecop()` when the margins have already been
transformed to approximately uniform pseudo-observations.

``` {r}
library(svines)
data(returns)  # data set of stock returns
returns <- returns[1:500, 1:2]
```

### Fitting models

```{r}
fit <- svine(returns, p = 1)  # Markov order 1
summary(fit)
```

```{r}
contour(fit$copula)
```

### Simulation

`svine_sim()` can be used in two different ways:

#### Generate a new time series of length 500
```{r}
sim <- svine_sim(n = 500, rep = 1, model = fit)
pairs(sim)
pairs(returns)
```

#### Sample conditionally on the observed past

```{r}
sim <- svine_sim(n = 1, rep = 100, model = fit, past = returns)
pairs(t(sim[1, , ]))
```


### Standard errors

To generate bootstrap replicates with the one-step block multiplier bootstrap,
use
```{r}
set.seed(2026)
models <- svine_bootstrap_models(2, fit)
summary(models[[1]])
```

### Discrete variables

Declare discrete variables through `var_types` and restrict their marginal
families to suitable discrete distributions. The following model uses Poisson
margins and a Gaussian pair-copula family.

```{r}
counts <- cbind(
  claims = rpois(250, lambda = 2),
  events = rpois(250, lambda = 4)
)
fit_discrete <- svine(
  counts,
  p = 1,
  var_types = c("d", "d"),
  margin_families = "pois",
  family_set = "gaussian"
)
fit_discrete
svine_sim(5, rep = 1, model = fit_discrete)
```

`svine()` constructs the required CDF and left-limit CDF values automatically.
Users calling `svinecop()` directly must supply all regular `F(x)` columns,
followed by one `F(x-)` column for each discrete variable.

## References

Nagler, T., Krüger, D., and Min, A. (2022). Stationary vine copula models for multivariate time series. *Journal of Econometrics, 227(2), pp. 305-324* [[pdf]](https://arxiv.org/abs/2008.05990) [[doi]](https://doi.org/10.1016/j.jeconom.2021.11.015)
